### 2018-02-20 Maths Notes/Mechanics/M2/Kinematics/Kinematics.tex
M2 Notes Update - Exam Questions

>>>>

\underline{\Huge Problems with Vectors}
\includegraphics[width=10cm]{vector_problems.png}\\
\textit{A ball B of mass 0.4 kg is struck by a bat at a point O which is 1.2 m above horizontal ground.
The unit vectors i and j are respectively horizontal and vertical. Immediately before being
struck, B has velocity $(-20i+4j) \ ms^{–1}$. Immediately after being struck it has velocity $(15i+
16j) \ ms^{–1}$.\\
After B has been struck, it moves freely under gravity and strikes the ground at the point A, as
shown in the diagram above. The ball is modelled as a particle.}\\
\textit{Calculate the magnitude of the impulse exerted by the bat on B. }
\textcolor{red}{$$I=m(v-u)$$}
Calculate the impulse as a vector using information from the question
$$I=0.4(15i+16j-(-20i+4j)=14i+4.8j$$
Calculate the magnitude of this vector
$$|I|=\sqrt{14^2+4.8^2}=14.5Ns$$
\textit{By using the principle of conservation of energy, or otherwise, find the speed of B when
it reaches A.}
Initial energy:
$$\frac{1}{2}\times0.4\times\Big(\sqrt{15^2+16^2}\Big)^2=96.2J$$
Energy gained from loss in GPE
$$0.4\times9.8\times1.2=4.704J$$
Add energies and set equal to final kinetic energy
$$4.704+96.2=\frac{1}{2}\times0.4\times v^2$$
Find v
$$v=\sqrt{\frac{4.704+96.2}{0.5\times0.4}}=22.46$$
\textit{Calculate the angle which the velocity of B makes with the ground when B reaches A.}\\
Using the knowledge that horizontal velocity is constant and the speed is known, use cos to find the angle
$$\arccos\Bigg(\frac{15}{22.5}\Bigg)=48\degree$$
\textit{State two additional physical factors which could be taken into account in a refinement of
the model of the situation which would make it more realistic}
\item Air resistance
\item Wind
\item Rotation of ball (ball is not a particle)

### 2018-02-20 Maths Notes/Mechanics/M2/Statics of rigid bodies/Statics of rigid bodies.tex
M2 Notes Update - Exam Questions

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}

@@

\begin{document}
\begin{center}
\underline{\huge Statics of rigid bodies}
\end{center}
\section{Equilibrium of rigid bodies}
When calculating moments, only use the component of the force acting perpendicular to the rod.\\
\\
A rigid body is in equilibrium if:
\begin{itemize}
\item The vector sum of the forces is zero
\item The sum of the moments around any point is zero
\end{itemize}

\end{document}

>>>>

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}

@@

\begin{document}
\begin{center}
\underline{\huge Statics of rigid bodies}
\end{center}
\section{Equilibrium of rigid bodies}
When calculating moments, only use the component of the force acting perpendicular to the rod.\\
\\
A rigid body is in equilibrium if:
\begin{itemize}
\item The vector sum of the forces is zero
\item The sum of the moments around any point is zero
\end{itemize}
\newpage
\begin{center}
\underline{\Huge Equilibrium of rigid bodies example}
\end{center}
\includegraphics[width=8cm]{eq_rigid.png}\\
\textit{The diagram above shows a uniform rod AB of mass m and length 4a. The end A of the rod is
freely hinged to a point on a vertical wall. A particle of mass m is attached to the rod at B. One
end of a light inextensible string is attached to the rod at C, where AC = 3a. The other end of the
string is attached to the wall at D, where AD = 2a and D is vertically above A. The rod rests
horizontally in equilibrium in a vertical plane perpendicular to the wall and the tension in the
string is T. }\\
\\
\textit{Show that T = $mg\sqrt{13}$}\\
\\
Take moments about A 
$$3a\times T\cos\theta=2amg+4amg$$
Calculate $\cos\theta$ from the lengths on the diagram
$$\cos\theta=\frac{2}{\sqrt{3^2+2^2}}=\frac{2}{\sqrt{13}}$$
Substitute in the value of $\cos\theta$ and divide through by $a$
$$\frac{6}{\sqrt{13}}T=6mg$$
Multiply both sides by $\frac{\sqrt{13}}{6}$
$$T=mg\sqrt{13}$$
\\
\textit{The particle of mass m at B is removed from the rod and replaced by a particle of mass M which
is attached to the rod at B. The string breaks if the tension exceeds $2mg\sqrt{13}$. Given that the
string does not break,\\
show that $M\leqslant\frac{5}{2}m$}\\
\\
Rewrite the moments equation with the new information from the question
$$3a\times T\cos\theta=2amg+4aMg$$
Write the inequality given
$$T\leqslant 2mg\sqrt{13}$$
Substitute in the value for T
$$\frac{2mg+4Mg}{6}\sqrt{13}\leqslant2mg\sqrt{13}$$
Simplify
$$mg+2Mg\leqslant 6mg$$
$$M\leqslant\frac{5}{2}m$$
\end{document}
